Deep Blockchains

### 2018-07-07 paper/main.tex
added smt package, added paper feedback

\title{Deep Blockchains}
\author{Sourabh Niyogi\\Wolk Inc.\\{\tt sourabh@wolk.com}}
\date{July 4, 2018}

\usepackage{url}
\usepackage{amsfonts}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[latin1]{inputenc}
\usepackage{tikz}
\usepackage[nottoc]{tocbibind}
\geometry{left=.75in,top=.75in,right=.75in,bottom=1.0in}

@@

Layer 1 blockchains such as Ethereum and Bitcoin, on their own, cannot support the latency and throughput needs for modern decentralized web applications.  Attempting to support higher throughput or lower latency with naive solutions (e.g. larger blocks, lower security consensus algorithms, etc.) sacrifices the core benefits of layer 1 blockchains.   It is unnecessary to make this sacrifice in the name of scalability for blockchain databases:  when one blockchain is capable of storing and retrieving state, then another blockchain's summary state variables may be stored there.  This can be done in layers, where {\em Layer i+1} blockchain's state is stored in {\em Layer i} blockchains and each blockchain uses a well-motivated consensus engine to achieve Byzantine fault tolerance.
Using this layered approach, we can specify the key elements of a {\em deep blockchain} architecture.  The blockchain paradigm \cite{wood2015} that forms the backbone of all decentralized consensus-based transaction systems to date is as follows. A valid state transition for a blockchain of Layer $i$ is one which comes about through a transaction $\transaction{i}{j}$:

@@

In a {\bf deep blockchain} system, we will say that blockchain layer $i$ is said to be {\em connected} to layer $i+1$ if:
\begin{enumerate}
    \item there exists a transaction mapping function $\submission{i+1}$ mapping blocks at layer $i+1$ into transactions $\transaction{i}{k}$ at layer $i$ for all blocks $\block{i+1}{j}$
\begin{equation}
\transaction{i}{k} \equiv \submission{i+1}(\block{i+1}{j})
\end{equation}
    \item there exists a mapping function $\retrieval{i}(k)$ retrieving from blockchain layer $i$ a mapping  $f(\block{i+1}{k})$ of the blocks state of layer $i+1$  for all blocks $\block{i+1}{k}$
\begin{equation}
\retrieval{i}(k) \equiv f(\block{i+1}{k})
\end{equation}
\end{enumerate}
A natural choice for transaction mapping $\submission{i+1}(\block{i+1}{j})$ may be to include a block hash $\blockhash{i+1}{k}$ of the block $\block{i+1}{k}$ as a transaction $\transaction{i}{k}$, and for the lower layer to provide the block hash back (see Figure \ref{timeline} (left)).  In this paper, we demonstrate a deep blockchain system for provable storage, situating a ``Plasma Cash'' design \cite{plasmacashsimplespec} in a Layer 2 Blockchain and NoSQL/SQL/File Storage for any number of Layer 3 Blockchains (see Figure \ref{deepblockchain}).

@@

Historically, the low-throughput high-latency of Layer 1 blockchains resulted in immediate pressure to drive activities off-chain, but
only a few ``off-chain'' attempts can be considered a deep blockchain because they lack the connected blockchains.  Layer $i+1$ and layer $i$  may be explicitly connected in a deep blockchain system for many different reasons:
\begin{enumerate}
    \item Higher throughput services at layer $i+1$ may be paid for using the value held in layer $i$ currency

@@

Seminal insights on multi-layer blockchains were put forth by \cite{poonbuterin2017}, which have inspired many ``Plasma'' designs, and specifically motivated our implementation of what has been termed ``Plasma Cash'' for tracking bandwidth balances.
The Layer 2 Plasma Cash blockchain is connected to Layer 1 using the following trust primitives:
\begin{itemize}
\item User Deposit: When Alice wishes to use the services enabled by the Layer 2 blockchain, Alice deposits some Layer 1 currency $\lambda_{dep}$ (.01 ETH or 1 WLK) in a Layer 1 contract function (\texttt{createBlockchain}); the deposit event results in Alice owning a Layer 2 token $\tau$ through a Layer 2 Deposit transaction included in a Layer 2 block.

@@

\item Layer 2 Block Connection: The operators of the Layer 2 blockchain mints new Layer 2 blocks $\block{2}{j}$ with a collation of Layer 2 transactions $\transaction{2}{k}$ (with a consensus protocol such as Quorum RAFT, POA) from the User Deposit and User Transfer transactions.  Each Layer 2 block $\block{2}{k}$ has its Merkle Root $\blockhash{2}{k}$ submitted to Layer 1 with a transaction $\transaction{2}{k} = \submitblock{k}{\blockhash{2}{k}}$ recorded in a Layer 1 block $\block{1}{l}$.   The recipient of a token transfer must receive the full history of all transactions from Alice and verify it against these Merkle Roots $\blockhash{2}{k}$ stored in Layer 1, all the way to the original deposit.

\item User Exit: When Alice wishes to withdraw her token $\tau$ for Layer 1 cryptocurrency, she calls $\startexit$ function with the last 2 transactions which can be verified against and Merkle proofs that must match the stored Merkle roots to be a valid exit; if no one challenges the exit, Alice receives the original $\lambda_{dep}$ within a short time period when exits are finalized.

@@

The specific deep blockchain system that we have developed extends the Blockchain 2.0 Objective up one more layer by incorporating trust primitives (Block transactions, Sparse Merkle Trees) in provable NoSQL, SQL and Storage services, shown in Figure 2:  Layer 3 NoSQL, SQL and Storage blockchains rest on the storage and bandwidth services of Layer 2, which supervene on the decentralized computation and payment services of Layer 1.  Our work follows Ethereum SWARM's foundational work on storage and bandwidth \cite{swarm} which outlines the following ideas that we situate in multiple layers:
\begin{itemize}
    \item Chunks are stored in Cloudstore by the hash of their content.  Nodes that request a chunk by key $k$ can verify correctness of the value $v$ returned by checking the hash of $v$.
    \item Insurers of chunks can earn Layer 1 currency with valid Merkle proofs; Failure to provide valid proofs result in severe insurance payouts
    \item Bandwidth consumed by a node, when hitting the nodes threshold must result in signed payments
\end{itemize}
Layer 1 blockchains were initially developed without the concern for storage models being competitive with cloud computing platforms or even a passing concern for bandwidth; the birth of Bitcoin and Ethereum Layer 1 focused on birthing trustless payments and trustless computation mediated by a peer-to-peer network, rather than about nodes providing decentralized storage.   In contrast, decentralized {\em storage} networks, as manifested in Ethereum SWARM and many other systems, promises to have a large peer-to-peer network of nodes sharing the responsibility to keep a portion of the world's data and compensated proportionately for the commodity storage and bandwidth they provide.  In these networks, a distributed hash table (typically, with Kademlia routing layers) is used for logarithmic look ups of chunks, but in practice, $O(log_2(n))$ retrieval times are just not competitive with modern UI expectations or typical developer expectations.  Nevertheless, decentralized storage networks have a critical role to play in providing censorship-resistance.  Rather than layer 3 rest solely on a decentralized storage network (which is slow but resilient and censorship-resistant), we can rest layer 3 on {\em both} decentralized storage networks {\em and} mature modern cloud computing platforms.  Again, the Blockchain 1.0 Objective must be displaced in favor of the Blockchain 2.0 Objective: in this sense, more storage variety {\em increases} the cost of attack.

Putting the elements together in a deep blockchain system for provable storage:
\begin{itemize}
\item Layer 1 blockchain: When a developer wishes to have a Layer 3 blockchain for NoSQL/SQL/Storage, they send Layer 1 currency into \texttt{createBlockchain(blockchainName string)} on MainNet; this can be refunded with a {\tt dropBlockchain(blockchainName string)} operation [taking place of {\tt startExit}].  When storage is used in {\tt blockchainName} through the activities of Layer 3 blockchains (as recorded by the Layer 2 blockchain below), this balance goes down.

\item Layer 2 Plasma Cash Blockchain: The storage and retrieval of chunks in Cloudstore are exposed to Layer 3 blockchains with the following 2 APIs (see Appendix A):
\begin{itemize}
    \item \texttt{storeChunk}$(k, v, \tau, \signature)$ - stores a key-value pair mapping $(k,v)$ in Cloudstore, backed by Layer 2 token $\tau$ (signed with $\signature$) received from the Layer 1 transaction.
    \item \texttt{retrieveChunk}$(k, \tau, \signature)$ - retrieves a key-value pair mapping $(k,v)$ in Cloudstore, backed by Layer 2 token $\tau$ (again, signed with $\signature$), and returning the balance of $\tau$ used so far
\end{itemize}
The Layer 2 operator will store via Cloudstore in as many regions and cloud providers as necessary to {\bf insure} the chunk as follows:  A new type of Layer 2 {\bf block transaction} insures a set of chunks recorded through \texttt{storeChunk} calls.  The cause of these chunks are from any Layer 3 blockchain needing storage and bandwidth, where bandwidth is used in \texttt{retrieveChunk} calls.  When a Layer 3 blockchain mints Layer 3 blocks, the Layer 3 blocks themselves contain a list of chunks.  The blocks themselves are submitted through \texttt{storeChunk} calls, signed by the Layer 3 blockchain owner, and the block hash $\blockhash{3}{k}$ is submitted by the Layer 3 blockchain to the Layer 2 blockchain via a $\submitblock{k}{\blockhash{3}{k}}$ block transaction.  This enables the Layer 2 blockchain to meter the cumulative storage of {\tt blockchainName} and deduct from the balance originally deposited in the $\texttt{createBlockchain(blockchainName string)}$ operation (approximately every 24 hours), passing on Cloudstore fiat and crypto costs to Layer 3 blockchains.  Notably, Layer 3 blocks themselves are recorded with $\texttt{storeChunk}(k, v, \tau, \signature)$ to store the layer 3 block in Cloudstore and then results in a call to $\submitblock{j}{\blockhash{3}{j}}$:
\begin{equation}
\transaction{3}{j} \equiv \submitblock{j}{\blockhash{3}{j}}
\end{equation}
Because the block storage is signed and because the block transactions are signed, Layer 2 operators  collect storage payments with the layer 3 blockchain operator consent, forming a kind of ``state channel'' within the deep blockchain.  Taken together, this is the Layer 3 Block Connection, as seen in Figure \ref{timeline}.  The Layer 2 block consists of:
\begin{itemize}
\item the transaction root $\transactionroot{2}{k}$ that utilizes the SMT structure to represent just the tokens $\tau_1, \tau_2, \ldots$ spent in block $k$
\begin{equation}
\transactionroot{2}{k} \equiv \texttt{KT}( (\tau_1, \transaction{2}{\tau_1}), (\tau_2, \transaction{2}{\tau_2}), ... )
\end{equation}
\item the token root $\tokenroot{k}$  for {\em all} tokens $\tau_j, ...$
\begin{equation}
\tokenroot{k} \equiv \texttt{KT}( (\tau_1, \transaction{2}{\tau_1}), (\tau_2, \transaction{2}{\tau_2}), ... )
\end{equation}
\item array of token transactions  $\transaction{2}{k}$
\item array of block transactions  $\blocktransaction{2}{k}$ from all Layer 3 blockchain operators using Layer 2 services
\item an account root, using an SMT to store an accounts ``balance'' and a list of tokens held by that account.
\end{itemize}

\item Layer 3 blockchains:  Any number of Layer 3 blockchains utilize storage and bandwidth can be layered on top of the Layer 2 blockchain, regularly submitting lists of chunks based on the structure of the Layer 3 blockchain.
\begin{itemize}
    \item For NoSQL + File Storage, there is a key for each row of NoSQL or File, and a value for the row (a JSON record) or raw file contents.   The root hash changes when any table schema is updated, and where each table has a root hash that changes when any record of the table is changed; any new database content results in new chunks, where the chunk is referenced by the hash of its content.
    \item For SQL, there is a root hash for each database, where the root hash changes when any table schema is updated, and where each table has a root hash that changes when any record of the table is changed; any new database content results in new chunks, where the chunk is referenced by the hash of its content.
\end{itemize}
We develop both NoSQL and SQL Blockchains in Section ~\ref{sec:layer3}.
\end{itemize}

\noindent Just as with Layer 1 blockchain nodes, running Layer 2 and Layer 3 blockchains consists of running a node in a peer-to-peer network, receiving and transmitting messages about new transactions and new blocks.   Our blockchain implementations of the Layer 2 and Layer 3 originated from the go-ethereum and JPMorgan Quorum RAFT code bases, written in Golang.  We used RAFT for both Layer 2 and Layer 3 implementations due to its simple model of finality.  For each blockchain, we created a Golang package containing each of the interfaces specified in Appendix A, and adapted Quorum RAFT code to conform to these interfaces.  There is no explicit assumption that  permissioned consensus algorithms be used, however.   The choice of RAFT was made purely out of simplicity, its maturity as a code base, and its capacity for high throughput -- any consensus protocol that achieves finality can fit within this deep blockchain architecture.  For both the Layer 2 and Layer 3 blockchains, we use a Sparse Merkle Tree to support provable data storage, which we review in detail next.

@@

With the foundations of Layer 2 providing storage and bandwidth paid for with Layer 2 tokens, any number of Layer 3 blockchains may be constructed.  We detail the construction of a NoSQL and SQL blockchain here.  Layer 3 blockchains record their block hashes in Layer 2, and because Layer 2 stores its block hashes in Layer 1 in the following way:
\begin{itemize}
 \item Each SQL and NoSQL transaction is included in a Level 3 block
 \item Each Level 3 block has its hash included in a Level 2 block
 \item Each Level 2 block has its Merkle Root included in {\tt submitBlock} recorded in a Level 1 block
\end{itemize}
It then becomes possible to aggregate multiple proof of inclusions at the highest layers all the way to MainNet with Deep Merkle Proofs, which we illustrate here.

@@

Let us illustrate a detailed example, shown in Figure~\ref{deepmerkleproof}:
\begin{itemize}
\item In Layer 3 Block 302, the user wishes store document ID 1 with key $K_1$ mapped to encrypted value $V_1$ and document ID 2 mapped to encrypted value $V_2$. The user can submit 2 Layer 3 NoSQL transactions:
\begin{equation*}
\left[ \texttt{SetKey}(K_1=\texttt{0b001...00},V_1=\texttt{0x778899..}),
\texttt{SetKey}(K_2, V_2)
\right]
\end{equation*}
\noindent which results in a set of SMT primitive operations:
\begin{equation*}
\{\texttt{insert}(\hash{K_1},\hash{\encrypt{V_1}}),  \texttt{insert}(\hash{K_2},\hash{\encrypt{V_2}}) \}
\end{equation*}
resulting in
$\texttt{Layer3KeyRoot}=\texttt{0x83fc...}$.  The chunks for both documents $H(V_1)$ and $H(V_2)$ along with chunk of the previous block 301  (e.b. \texttt{storeChunk(0b001.., ...)}) are included in Layer 3 Block 302 in the \texttt{ChunkIDs} and insured with a call to \[
\texttt{submitBlock(0b101..11, 302}
\]
submitted to the Layer 2 blockchain.

@@

\item An array of SQL transactions that is mapped into newly created chunks (created via $\texttt{storeChunk}$ for table root hashes

@@

\item When an owner creates a new database, the owner chunk is updated and  database chunk is created and the owner chunk is updated with the new database chunk information. If this is the first database created by the owner, the root hash of the owner is set for the first time. The root hash of the database is set for the first time.
\item When an owner creates a new table, the database chunk is updated and table chunk is created and the database chunk is updated with the new table information.  This also causes the owner chunk to be updated with the new database chunk information.  The root hash of the table is set for the first time in the child chain.
\item When an owner creates or updates a table, this creates or changes the database schema chunk.  The database chunk is then updated with the new schema information, which in turn causes the owner chunk to be updated with the new database chunk information.  (Describe the updating of the chunk with the database schema, something about a root hash and child chain here?)

@@

\item the ChunkIDs associated with the SQL operations, where chunks are only created, and never ``updated''.

@@

\section{Paying for Storage and Bandwidth}

The Layer 3 blockchain users who store NoSQL/SQL/File data with \texttt{storeChunk} operations give permissions to be charged for bandwidth and storage in two different ways:

\begin{enumerate}
\item {\em Bandwidth} is paid for through (1) signed  \texttt{retrieveChunk}$(k, \tau, \signature)$ calls to retrieve data and obtain recent balances, where each call uses up a tiny amount of bandwidth backed for with a token $\tau$ originated by the \texttt{createBlockchain} call; (2) signed \texttt{updateBalance} calls to agree to recent balances

\item {\em Storage} is paid for through Block Transactions $\submitblock{k}{\blockhash{3}{k}}$ signed by the Layer 3 block operator - because chunks are identified directly inside Layer 3 blocks, a tally of the number of bytes used in each new layer 3 block is added to the SMT.
\end{enumerate}
In this way Layer 3 Blockchains pay for the services of the Layer 2 blockchain.  The lifecycle of a short lived Layer 3 blockchain is shown in Figure~\ref{statechannel}, which we expound in the next section.

>>>>

\title{Deep Blockchains}
\author{Sourabh Niyogi, Michael Chung and Rodney Witcher\\Wolk Inc.\\
{\tt \{sourabh,michael,rodney\}@wolk.com}}
\date{July 7, 2018}

\usepackage{url}
\usepackage{amsfonts}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[latin1]{inputenc}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage[nottoc]{tocbibind}
\geometry{left=.75in,top=.75in,right=.75in,bottom=1.0in}

@@

Layer 1 blockchains such as Ethereum and Bitcoin, on their own, cannot support the latency and throughput needs for modern web applications.  Attempting to support higher throughput or lower latency with naive solutions (e.g. larger blocks, lower security consensus algorithms, etc.) sacrifices the core benefits of layer 1 blockchains.   It is unnecessary to make these sacrifices in the name of scalability for blockchains:  when one blockchain is capable of storing and retrieving state, then another blockchain's summary state variables may be stored there.  This can be done in layers, where {\em Layer i+1} blockchain's state is stored in {\em Layer i} blockchains and each blockchain uses a well-motivated consensus engine to achieve Byzantine fault tolerance.
Using this layered approach, we can specify the key elements of a {\em deep blockchain} architecture.  The blockchain paradigm \cite{wood2015} that forms the backbone of all decentralized consensus-based transaction systems to date is as follows. A valid state transition for a blockchain of Layer $i$ is one which comes about through a transaction $\transaction{i}{j}$:

@@

In a {\bf deep blockchain} system, we will say that blockchain layer $i$ is said to be {\em connected} to layer $i+1$ if:
\begin{enumerate}
    \item there exists a transaction mapping function $\submission{i+1}$ mapping blocks at layer $i+1$ into transactions $\transaction{i}{k}$ at layer $i$ for all layer $i+1$ blocks $\block{i+1}{j}$
\begin{equation}
\transaction{i}{k} \equiv \submission{i+1}(\block{i+1}{j})
\end{equation}
    \item there exists a mapping function $\retrieval{i}(k)$ retrieving from blockchain layer $i$ a mapping  $f(\block{i+1}{k})$ of the blocks state of layer $i+1$  for all blocks $\block{i+1}{k}$
\begin{equation}
\retrieval{i}(k) \equiv f(\block{i+1}{k})
\end{equation}
\end{enumerate}
A natural choice for transaction mapping $\submission{i+1}(\block{i+1}{j})$ may be to include a block hash $\blockhash{i+1}{k}$ of the block $\block{i+1}{k}$ as a transaction $\transaction{i}{k}$, and for the lower layer to provide the block hash back (see Figure \ref{timeline} (left)).  In this paper, we demonstrate a deep blockchain system for provable storage, situating a ``Plasma Cash'' design \cite{plasmacashsimplespec} in a Layer 2 Blockchain and NoSQL/SQL/File Storage for any number of Layer 3 Blockchains (see Figure \ref{deepblockchain}).

@@

Historically, the low-throughput high-latency of Layer 1 blockchains resulted in immediate pressure to drive activities off-chain, but
only a few ``off-chain'' attempts can be considered deep blockchains because they lack the connected blockchains.  Layer $i+1$ and layer $i$  may be explicitly connected in a deep blockchain system for many different reasons:
\begin{enumerate}
    \item Higher throughput services at layer $i+1$ may be paid for using the value held in layer $i$ currency

@@

Seminal insights on multi-layer blockchains were put forth by \cite{poonbuterin2017}, which have inspired many ``Plasma'' designs, and specifically motivated our implementation of what has been termed ``Plasma Cash'' for tracking storage and  bandwidth balances.  The Layer 2 Plasma Cash blockchain is connected to Layer 1 using the following trust primitives:
\begin{itemize}
\item User Deposit: When Alice wishes to use the services enabled by the Layer 2 blockchain, Alice deposits some Layer 1 currency $\lambda_{dep}$ (.01 ETH or 1 WLK) in a Layer 1 contract function (\texttt{createBlockchain}); the deposit event results in Alice owning a Layer 2 token $\tau$ through a Layer 2 Deposit transaction included in a Layer 2 block.

@@

\item Layer 2 Block Connection: The operators of the Layer 2 blockchain mints new Layer 2 blocks $\block{2}{j}$ with a collation of Layer 2 transactions $\transaction{2}{k}$ (with a consensus protocol such as Quorum RAFT and POA in permissioned networks or Ethereum Casper for permissionless networks) from the User Deposit and User Transfer transactions.  Each Layer 2 block $\block{2}{k}$ has its Merkle Root $\blockhash{2}{k}$ submitted to Layer 1 with a transaction $\transaction{2}{k} = \submitblock{k}{\blockhash{2}{k}}$ recorded in a Layer 1 block $\block{1}{l}$.   The recipient Bob of a token transfer must receive the full history of all transactions from Alice and verify it against these Merkle Roots $\blockhash{2}{k}$ stored in Layer 1, all the way to the original deposit.  If any transaction in the history cannot be verified by Bob, Bob cannot accept Alice's token as payment.

\item User Exit: When Alice wishes to withdraw her token $\tau$ for Layer 1 cryptocurrency, she calls $\startexit$ function with the last 2 transactions\footnote{As to why {\em two}, two is indicative, but not conclusive concerning Alice's ownership, therefore a user challenge process is required.} which can be verified against and Merkle proofs that must match the stored Merkle roots to be a valid exit; if no one challenges the exit, Alice receives the outstanding token balance within a short time period when exits are finalized.

@@

The specific deep blockchain system that we have developed extends the Blockchain 2.0 Objective up one more layer by incorporating trust primitives (Block transactions, Sparse Merkle Trees) in provable NoSQL, SQL and Storage services, shown in Figure 2:  Layer 3 NoSQL, SQL and Storage blockchains rest on the storage and bandwidth services of Layer 2, which supervene on the decentralized computation and payment services of Layer 1.  Our work follows Ethereum SWARM's foundational work on storage and bandwidth \cite{swarm} which outlines the following ideas that we situate in multiple layers:
\begin{itemize}
    \item A chunk of bytes $v$ is stored in Cloudstore using 256-bit hash $k = \hash{v}$ as the key to retrieve $v$.  Nodes that request a chunk by key $k$ can verify correctness of the value $v$ returned from Cloudstore simply by checking if   $k=\hash{v}$.
    \item Insurers of chunks can earn Layer 1 currency with valid Merkle proofs; Failure to provide valid proofs result in severe insurance payouts
    \item Bandwidth consumed by a node, when hitting the nodes threshold must result in signed payments
\end{itemize}
Layer 1 blockchains were initially developed without the concern for storage models being competitive with cloud computing platforms or even a passing concern for bandwidth; the birth of Bitcoin and Ethereum Layer 1 focused on birthing trustless payments and trustless computation mediated by a peer-to-peer network, rather than about nodes providing decentralized storage.   In contrast, decentralized {\em storage} networks, as manifested in Ethereum SWARM and many other systems, promises to have a large peer-to-peer network of nodes sharing the responsibility to keep a portion of the world's data and compensated proportionately for the commodity storage and bandwidth they provide.  In these networks, a distributed hash table (typically, with Kademlia routing layers) is used for logarithmic look ups of chunks, but in practice, $O(log_2(n))$ retrieval times are just not competitive with modern UI expectations or typical developer expectations.  Nevertheless, decentralized storage networks have a critical role to play in providing censorship-resistance.  Rather than layer 3 rest solely on a decentralized storage network (which is slow but resilient and censorship-resistant), we can rest layer 3 on {\em both} decentralized storage networks {\em and} mature modern cloud computing platforms.  Again, the Blockchain 1.0 Objective must be displaced in favor of the Blockchain 2.0 Objective: in this sense, more storage variety {\em increases} the cost of attack.

Putting the elements together in a deep blockchain system for provable storage:
\begin{itemize}
\item Layer 1 blockchain: When a developer wishes to have a Layer 3 blockchain for NoSQL/SQL/Storage, they send Layer 1 currency into \texttt{createBlockchain(blockchainName string)} on MainNet; this can be refunded with a {\tt dropBlockchain(blockchainName string)} operation (taking place of {\tt startExit}).  When storage is used in {\tt blockchainName} through the activities of Layer 3 blockchains (as recorded by the Layer 2 blockchain below), this balance goes down.  Balances can added to and withdrawn by the owner of the blockchain.

\item Layer 2 Plasma Cash Blockchain: The storage and retrieval of chunks in Cloudstore are exposed to Layer 3 blockchains with the following 2 APIs (see Appendix A):
\begin{itemize}
    \item \texttt{storeChunk}$(k, v, \tau, \signature)$ - stores a key-value pair mapping $(k,v)$ in Cloudstore, backed by Layer 2 token $\tau$ (signed with $\signature$) received from the Layer 1 transaction.
    \item \texttt{retrieveChunk}$(k, \tau, \signature)$ - retrieves a key-value pair mapping $(k,v)$ in Cloudstore, backed by Layer 2 token $\tau$ (again, signed with $\signature$), and returning the balance of $\tau$ used so far
\end{itemize}
The Layer 2 operator will store via Cloudstore in as many regions and cloud providers as necessary to {\bf insure} the chunk as follows:  A new type of Layer 2 {\bf block transaction} insures a set of chunks recorded through \texttt{storeChunk} calls.  The cause of these chunks are from any Layer 3 blockchain needing storage and bandwidth, where bandwidth is used in \texttt{retrieveChunk} calls.  When a Layer 3 blockchain mints Layer 3 blocks, the Layer 3 blocks themselves contain a Cloudstore key that references a list of chunks written in the Layer 3 block.  The block itself is stored in Cloudstore with another \texttt{storeChunk} call, signed by the Layer 3 blockchain owner, and the block hash $\blockhash{3}{k}$ is submitted by the Layer 3 blockchain to the Layer 2 blockchain via a $\submitblock{k}{\blockhash{3}{k}}$ block transaction.  This enables the Layer 2 blockchain to meter the cumulative storage of {\tt blockchainName} and deduct from the balance originally deposited in the $\texttt{createBlockchain(blockchainName string)}$ operation (approximately every 24 hours), passing on Cloudstore costs to Layer 3 blockchains.  Notably, Layer 3 blocks themselves are recorded with $\texttt{storeChunk}(k, v, \tau, \signature)$ to store the layer 3 block in Cloudstore and then results in a call to $\submitblock{j}{\blockhash{3}{j}}$:
\begin{equation}
\transaction{3}{j} \equiv \submitblock{j}{\blockhash{3}{j}}
\end{equation}
Because the block storage is signed and because the block transactions are signed, Layer 2 operators  collect storage payments with the layer 3 blockchain operator consent, forming a kind of ``state channel'' within the deep blockchain.  Taken together, this is the Layer 3 Block Connection, as seen in Figure \ref{timeline}.  The Layer 2 block consists of:
\begin{itemize}
\item the transaction root $\transactionroot{2}{k}$ that utilizes the SMT structure to represent just the tokens $\tau_1, \tau_2, \ldots$ spent in block $k$
\begin{equation}
\transactionroot{2}{k} \equiv \texttt{KT}( (\tau_1, \transaction{2}{\tau_1}), (\tau_2, \transaction{2}{\tau_2}), ... )
\end{equation}
\item the token root $\tokenroot{k}$  for {\em all} tokens $\tau_j, ...$
\begin{equation}
\tokenroot{k} \equiv \texttt{KT}( (\tau_1, \transaction{2}{\tau_1}), (\tau_2, \transaction{2}{\tau_2}), ... )
\end{equation}
\item array of token transactions  $\transaction{2}{k}$
\item array of block transactions  $\blocktransaction{2}{k}$ from all Layer 3 blockchain operators using Layer 2 services
\item an account root, using an SMT to store an accounts ``balance'' and a list of tokens held by that account.
\end{itemize}

\item Layer 3 blockchains:  Any number of Layer 3 blockchains that utilize storage and bandwidth can be layered on top of the Layer 2 blockchain, regularly submitting lists of chunks based on the structure of the Layer 3 blockchain.
\begin{itemize}
    \item For NoSQL + File Storage, there is a key for each row of NoSQL or File, and a value for the row (a JSON record) or raw file contents.   The root hash changes when any table is added/removed or when any table schema is updated, and where each table has a root hash that changes when any record of the table is changed; any new database content results in new chunks, where the chunk is referenced by the hash of its content.
    \item For SQL, there is a root hash for each database, where the root hash changes when any table schema is updated, and where each table has a root hash that changes when any record of the table is changed; any new database content results in new chunks, where the chunk is referenced by the hash of its content.
\end{itemize}
We describe both NoSQL and SQL Blockchains in Section ~\ref{sec:layer3}.
\end{itemize}

\noindent Just as with Layer 1 blockchain nodes, running Layer 2 and Layer 3 blockchains consists of running a node in a peer-to-peer network, receiving and transmitting messages about new transactions and new blocks.   Wolk's blockchain implementations of the Layer 2 and Layer 3 originated from Ethereum's \href{https://github.com/ethereum/go-ethereum}{go-ethereum} and JPMorgan's \href{https://github.com/jpmorganchase/quorum}{Quorum RAFT} code bases, written in Golang.  We used RAFT for both Layer 2 and Layer 3 implementations due to its simple model of finality.  For each blockchain, we created a Golang package containing each of the interfaces specified in Appendix A, and adapted Quorum RAFT code to conform to these interfaces.  There is no explicit assumption that  permissioned consensus algorithms be used, however.   The choice of RAFT was made purely out of simplicity, its maturity as a code base, and its capacity for high throughput -- any consensus protocol that achieves finality can fit within this deep blockchain architecture.  For both the Layer 2 and Layer 3 blockchains, we use a \href{https://github.com/wolkdb/deepblockchains/smt}{Sparse Merkle Tree} to support provable data storage, which we review in detail next.

@@

With the foundations of Layer 2 providing storage and bandwidth paid for with Layer 2 tokens, any number of Layer 3 blockchains may be constructed.  We detail the construction of a NoSQL and SQL blockchain here.  At a high level, Layer 3 blockchains collate SQL and NoSQL transactions in Layer 3 blocks submit block transactions to Layer 2, and Layer 2 collate token and block transactions with Merkle Roots of token root and blocks submitted and included in transactions to Layer 1 blockchain.  It then becomes possible to aggregate multiple proof of inclusions at the highest layers all the way to MainNet with Deep Merkle Proofs, which we illustrate here.

@@

We present a detailed example: (shown in Figure~\ref{deepmerkleproof})
\begin{itemize}
\item In Layer 3 Block 302, the user wishes store document ID 1 with key $K_1$ mapped to encrypted value $V_1$ and document ID 2 mapped to encrypted value $V_2$. The user can submit 2 Layer 3 NoSQL transactions:
\begin{equation*}
\left[ \texttt{SetKey}(K_1=\texttt{0b001...00},V_1=\texttt{0x778899..}),
\texttt{SetKey}(K_2, V_2)
\right]
\end{equation*}
\noindent which results in a set of SMT primitive operations:
\begin{equation*}
\{\texttt{insert}(\hash{K_1},\hash{\encrypt{V_1}}),  \texttt{insert}(\hash{K_2},\hash{\encrypt{V_2}}) \}
\end{equation*}
resulting in
$\texttt{Layer3KeyRoot}=\texttt{0x83fc...}$.  The chunks for both documents $H(V_1)$ and $H(V_2)$ along with chunk of the previous block 301  (e.b. \texttt{storeChunk(0b001.., ...)}) are included in Layer 3 Block 302 in the \texttt{ChunkIDs} and insured with a call to \[
\texttt{submitBlock(0b101..11, 302}
\]
submitted to the Layer 2 blockchain.

@@

\item An array of SQL transactions that is mapped into newly created chunks (created via $\texttt{storeChunk}$ for table root hashes)

@@

\item When the owner of a Layer 3 blockchain creates a new database, the owner chunk is updated and  database chunk is created and the owner chunk is updated with the new database chunk information. If this is the first database created by the owner, the root hash of the owner is set for the first time. The root hash of the database is set for the first time.
\item When the owner of a Layer 3 blockchain creates a new table, the database chunk is updated and table chunk is created and the database chunk is updated with the new table information.  This also causes the owner chunk to be updated with the new database chunk information.  The root hash of the table is set for the first time in the child chain.
\item When the owner of a Layer 3 blockchain creates or updates a table, this creates or changes the database schema chunk.  The database chunk is then updated with the new schema information, which in turn causes the owner chunk to be updated with the new database chunk information.

@@

\item the Chunks newly written through the execution of the SQL transactions, where chunks are only created, and never ``updated''.

@@

\section{Paying for Storage and Bandwidth}
The Layer 3 blockchain users who store NoSQL/SQL/File data with
\texttt{storeChunk} operations give the Layer 2 operator permission to charge for bandwidth and storage in two different ways:

\begin{enumerate}
\item {\em Bandwidth} is paid for through (1) signed  \texttt{retrieveChunk}$(k, \tau, \signature)$ calls to retrieve data and obtain recent balances, where each call uses up a tiny amount of bandwidth backed for with a token $\tau$ originated by the \texttt{createBlockchain} call; (2) signed a new \texttt{updateBalance} call to agree to recent balances;

\item {\em Storage} is paid for through Block Transactions $\submitblock{k}{\blockhash{3}{k}}$ signed by the Layer 3 blockchain operator - because chunks are identified directly inside Layer 3 blocks, a tally of the number of bytes used in each new layer 3 block is added to the SMT.  The Layer 1 contract then exposes \texttt{storageCharge} interface to the Layer 2 operator where a recently signed Layer 2 block transaction (containing a tally of the number of bytes, signed by the Layer 3 blockchain operator) is used to deduct the layer 3 operator's balance since the last time it was called.   This is detailed below.
\end{enumerate}
In this way Layer 3 Blockchains pay for the services of the Layer 2 blockchain.  The lifecycle of a short lived Layer 3 blockchain is shown in Figure~\ref{statechannel}, which we expound in the next section.
